---
title: "Untitled"
author: "Jake Eisaguirre"
date: "8/18/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)

#specifically for dataviz:
library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)

library(gapminder)

library(sf)
```


```{r}

liz <- read_csv(here("data_tidy", "lizards.csv"))

ggplot(liz, aes(x = weight)) +
  
  geom_histogram(fill = "orange", 
                 color = "blue",
                 size = .3,
                 linetype = "dotted") 
  


```
```{r}

ggplot(liz, aes(x = total_length, y = weight)) +
  geom_point(shape = 22,
             fill = "blue",
             color = "red",
             size = 4,
             alpha = 0.5)

```
bubble plot where the color of the points changes based on common name and the size of points varies based on total length. 
```{r}

ggplot(liz, aes(x = total_length, y = weight)) +
  
  geom_point(aes(color = common_name, size = total_length),
             shape = 23,
             fill = "black") +
  theme_classic()

```

```{r}


ggplot(liz, aes(x = total_length, y = weight)) +
  
  geom_point(aes(color = common_name)) +
  
  theme_classic() +
  
  facet_wrap(~common_name)



```

```{r}

ggplot(liz, aes(x = total_length, y = weight)) +
  
  geom_point(aes(color = common_name)) +
  
  facet_grid(sex ~ tail) +
  
  theme_classic()

```

```{r}

lizard_counts <- liz %>% 
  
  group_by(common_name) %>% 
  
  summarize(count = n())


lizard_counts <- liz %>% 
  
  count(common_name)

```


#
```{r}

#fct_reorder aligns by height

ggplot(lizard_counts, aes(x = count, y = fct_reorder(common_name, count))) +
  
  geom_col(aes(fill = common_name)) +
  
  theme_classic() +
  
  labs(x = "Lizard Counts",
       y = "")

```

Try converting common name to an ordered factor outside of ggplot. fct_reorder() <- good function

```{r}

lizards <- lizard_counts %>% 
  mutate(common_name = fct_reorder(common_name, count))

```

## Axis scale breaks, limite & labels 

scatterplot: total_length vs weight
```{r}

ggplot(liz, aes(x = total_length, y = weight)) +
  
  geom_point() +
  
  theme_classic() +
  
  scale_x_continuous(breaks = c(0, 50, 500),
                     limits = c(0, 500),
                     expand = c(0,0)) +
  
  scale_y_continuous(expand = c(0,0))

```

transrome the date column to class date, then find counts of observations by date
```{r}
library(lubridate)

liz_count <- liz %>% 
  
  mutate(date = mdy(date)) %>% 
  
  count(date)


ggplot(liz_count, aes(x = date, y = n)) +
  
  geom_line() +
  
  theme_classic() + 
  
  scale_x_date(date_breaks = "2 years",
               date_labels = "%y")
  
  
  

```

