---
title: "Untitled"
author: "Jake Eisaguirre"
date: "8/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
```

```{r}

liz <-  read_csv(here("data_tidy", "lizards.csv"))

```

## ggrepel

make a subset from lizards, called ww_liz, that only contains observatoins fo wwtail liz from the site "sand"
```{r}

ww_liz <- liz %>% 
  
  filter(common_name == "western whiptail", site == "sand")

ggplot(ww_liz, aes(x = total_length, y = weight)) +
  
  geom_point() +
  
  theme_classic() +
  
  geom_text_repel(aes(label = toe_num), size = 3, max.overlaps = 20)

```
create a scatter plot of gdp per cap (x) and lifeexp (y), labeled by country name
```{r}

data <- gapminder %>% 
  
  filter(continent == "Europe", year == 1952)

ggplot(data, aes(x = gdpPercap, y = lifeExp)) +
  
  geom_point() +
  
  geom_text_repel(aes(label = country), size =  3) +
  
  theme_classic()
  
  
  



```

## gghighlight

```{r}

p <- ggplot(liz, aes(x = total_length, y = weight)) +
  geom_point()

p + gghighlight(toe_num == 250, label_key = toe_num)


```

```{r}

q <- ggplot(liz, aes(x = total_length, y = weight)) +
  
  geom_line(aes(color = common_name)) +
  
  gghighlight(max(weight) > 30)

q

```

## patchwork for compound figures

```{r}

p | q

(p | q) / q & theme_classic()

```

## a few new graph types

### Marginal plots

```{r}

whiptails <- liz %>% 
  
  filter(common_name == "western whiptail") %>% 
  
  drop_na(total_length, weight)

ggplot(data = whiptails, aes(x = total_length, y = weight)) +
  
  geom_point() + 
  
  geom_rug()

```

marginal plot with box plots

```{r}

j <- ggplot(data = whiptails, aes(x = total_length, y = weight)) +
  
  geom_point(aes(color = sex), size = 2) + 
  
  scale_color_manual(values = c("cyan4", "black", "goldenrod"),
                     name = "Sex:",
                     labels = c("female", "juvenile", "male")) +
 
   theme_minimal() +
  
  theme(legend.position = "bottom")

ggMarginal(j, type = "boxplot", groupColour = T)

```

### Beesarm plot

```{r}

ggplot(whiptails, aes(x = sex, y = weight)) +
  
  geom_beeswarm() +
  
  geom_violin(fill = NA) +
  
  geom_boxplot(fill = NA)

```

### A heat map with geom_tile()

make a new data  fram called lizard_counts, starting from lizards, with date converted to class date, then count by year and common name the number of lizards observed
```{r}

lizard_counts <- liz %>% 
  
  mutate(date = mdy(date)) %>% 
  
  mutate(year = year(date)) %>% 
  
  count(year, common_name) %>% 
  
  rename(count = n)

```

```{r}

ggplot(data = lizard_counts, aes(x = year, y = common_name)) +
  
  theme_classic() +
  
  geom_tile(aes(fill = count)) +
  
  geom_text(aes(label = count), color = "white", size = 3) +
  
  scale_fill_gradientn(colors = c("navy", "red", "orange")) + 
  
  scale_fill_viridis_c()

```
### make a map!
```{r}

jornada_veg <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml")) %>% 
  
  select(Name) %>% 
  
  clean_names()

ggplot(data = jornada_veg) +
  
  geom_sf(aes(fill = name), color = NA) +
  
  scale_fill_viridis_d() +
  
  theme_minimal() +
  
  labs(x =  "Longitude",
       y = "Latitude",
       fill = "Dominant vegetation:")

```

